### 2024-04-22 paper/paper.qmd
results updated

---
title: "Stream of Consciousness Literature: A 'Joyceless' Linguistic Landscape" 
subtitle: "Exploring Word Frequency, Sentiment Value and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg"
author: Quang Mai
thanks: "Code and data are available at: https://github.com/ponolite/stream-consciousness-language.git"
date: today
date-format: long
abstract: "This project focuses on understanding the language used by renowned stream of consciousness (SOC) authors James Joyce, Virginia Woolf, Marcel Proust, Katherine Mansfield and T.S Eliot. By conducting word frequency analysis and sentiment analysis on these authors' nine novels, this paper aims to uncover shared linguistic patterns and gain insights into the authors' mental states. This paper attempts to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society and its literary circle from late 19th to mid-20th century. (add one sentence on main results)"
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(tidytext)
library(rmarkdown)
library(gutenbergr)
library(tm)
library(dplyr)
library(ggplot2)
library(scales)
library(knitr)
library(igraph)
library(widyr)
library(ggraph)
library(textdata)
library(arrow)
library(kableExtra)

portrait_clean <- read_parquet(here::here("data/analysis_data/portrait_clean.parquet"))
portrait_words <- read_parquet(here::here("data/analysis_data/portrait_words.parquet"))
swann_words <- read_parquet(here::here("data/analysis_data/swann_words.parquet"))
dalloway_words <- read_parquet(here::here("data/analysis_data/dalloway_words.parquet"))
bliss_words <- read_parquet(here::here("data/analysis_data/bliss_words.parquet"))
prufrock_words <- read_parquet(here::here("data/analysis_data/prufrock_words.parquet"))
combined_text <- read_parquet(here::here("data/analysis_data/combined_text.parquet"))
portrait <- read_csv(here::here("data/raw_data/portrait.csv"))
all_books <- read_csv(here::here("data/raw_data/all_books.csv"))
combined_books <- read_csv(here::here("data/raw_data/combined_books.csv"))

```

# Introduction

Stream of consciousness (SOC) is a narrative technique that aims to capture the continuous flow of thoughts, feelings, and sensations experienced by a character without conventional organization or punctuation [@resampling]. It mirrors the unpredictable and interconnected nature of human thought processes, often revealing the inner workings of the character's mind in an intimate and unfiltered manner [@flow]. In literature, most scholars agree that stream of consciousness reveals the complexities of mind-scapes, shedding light on the nuances of characters' emotional well-being and psychological struggles [@pathology]. As such, this paper has mined the texts of a total of nine novels from the volunteer archive, Project Gutenberg, to examine the mental health themes of famous stream of consciousness authors, namely by Joyce, Woolf, Proust, Mansfield and Eliot, from the modernist era of literature, spanning from late 19th century to mid-20th century [@gutenberg].

By analyzing these textual datasets through word frequency and sentiment analysis, I seek to pose and answer crucial questions: \textit{What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness? Moreover, how does this relationship vary differently across differnet demographics of authors, for instance, authors with different geographical locations and genders?} Understanding these dynamics is crucial in having an informed understanding of the West's late 19th to mid-20th century literature and even socio-political landscape, especially in regards to how authors and creative writers navigate and deal with then-taboo topics such as existential angst, mental health issues and disabilities.

Thus, the estimand is the correlation between mental health-related themes and words in SOC literature, their frequency and sentiment value as provided by @rTidytext. This is considered in terms of nine selected SOC novels only, namely Joyce's \textit{A Portrait of the Artist as a Young Man and Chamber Music}; Woolf's \textit{Mrs Dalloway and Jacob's Room}; Proust's \textit{Swann Way}; Mansfield's \textit{Bliss and The Garden Party}; and Eliot's \textit{The Waste Land and The Love Song of J. Alfred Prufrock}. Through our analysis, we found that (percentage, number and data here, main results)...

To further understand the correlation between stream of consciousness novels and mental health themes, in [Introduction], the paper briefly discusses the nature of stream of consciousness literature, relevant authors and the works that I've chosen to analyze. Subsequently, in [Data] and [Results], I talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts. Next, [Discussion] provides further insights and future areas of study. Finally, [Conclusion] summarizes our main findings. To complete the paper, [Appendix] clarifies how each variable within each dataset is generated with relevant tables to accordingly demonstrate this.

The novel texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rGuten] [@gutenberg]. Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg], `textdata` [@rTextdata], `tm` [@rTm], `here` [@here], `kableExtra` [@kableExtra], `arrow` [@arrow], and `knitr` [@rKnitr].

# Data {#sec-data}

## Measurement

Two central variables in this paper are:

* **Word Frequency:** This variable captures the repetition of a single word (unigram), two-words combination (bigram) or three-words combination (trigram) throughout a SOC novel text, providing us with a thematic understanding of SOC literature.
* **Sentiment Value:** This variable enables us to analyze how every word is usually perceived emotionally, whether it be a qualitative feeling or if it is conveyed through a numerical value.

Out of two variables used, the first one, 'Word Frequency' usually captured as `n` or `frequency` in datasets, is directly quantified through tokenizing the novel texts using the package `tidytext` and its function `unnest_tokens()` [@rTidytext]. To do this, I first downloaded all nine novel texts from @gutenberg, and leveraged functions such as `unnest_tokens()` from @rTidytext to mine the texts, or separating it into individual words. Finally, I used `count()` to quantify the word frequency.

The second variable used, 'Sentiment Value', is based on three English-based, general-purpose  "word-emotion and word-polarity association lexicons", sourced from Mohammad and Turney's expansive research along with efforts from Finn Årup Nielsen and Bing Liu and collaborators [@rTidytext] [@sentiment]. The three general-purpose lexicon that contributes to my sentiment analysis are [@rTidytext]:

- 'AFINN' from Finn Årup Nielsen, which assigns each word a numerical value from '-5 to 5', judging its emotionality from negative to postive
- 'bing' from Bing Liu and collaborators, which assigns if a word is either 'positive' or negative'
- 'nrc' from Saif Mohammad and Peter Turney, which assigns a core emotional value to a word, such as 'fear', 'anger', 'sadness' or 'trust'.

In terms of measuring 'Sentiment Value', all three general-purpose lexicons are compiled through crowd-sourcing and directly surveying the public on how each word is emotionally perceived. A survey sample of how the word 'startle' is compiled within the 'nrc' lexicon is presented below [@sentiment]:

(@) Which word is closest in meaning (most related) to startle?
- automobile
- shake
- honesty
- entertain

(@) How positive (good, praising) is the word startle?:
- startle is not positive
- startle is weakly positive
- startle is moderately positive
- startle is strongly positive

(@) How negative (bad, criticizing) is the word startle?
- startle is not negative
- startle is weakly negative
- startle is moderately negative
- startle is strongly negative

After the survey results are garnered, researchers average the answers to sort each surveyed word into pre-defined categories, specifically '-5 to 5' for 'AFINN', 'postive' or 'negative' for 'bing', and 'anger' or 'fear' for 'nrc'. With the continuous work of compiling these lexicons spanning years and decades [@sentiment] comes these functions: `get_sentiments("bing")` and `get_sentiments("nrc")` and `get_sentiments("afinn")` in which I can use `inner_join` to categorize my existing datasets of novel texts into their sentiment value. Systematic and data-driven, these measurement methods ensure that all lexicons faithfully reflects each word's emotionality.

However, I do recognize how reductive this quantification of language can be. When it comes to understand such social and human artifacts as language or literary texts, much is dependent on their context. As such, I will further discuss these weaknesses of the datasets under [Discussion].

## Source Data: Project Gutenberg

Founded in December 1, 1971 by Michael S. Hart, Project Gutenberg exclusively publishes literature in the public domain within the United States. Typically, submissions to Project Gutenberg are digitized editions of printed books, the majority of which were published over 95 years ago. To confirm public domain status, authors can use the copy.pglaf.org website [@gutenberg].

The archive relies heavily on volunteers for support and content selection, but it does not accept copyrighted or modern works, even with permission. To submit work, individuals must obtain copyright clearance, scan or capture book images, dedicate hours to proofreading and formatting, and ensure compliance with Project Gutenberg’s guidelines, including valid HTML and correct spelling. Common points of failures for submitted works include:

- Crop marks or other printer's marks should not be used on any files.
- The book description should be in compliance with the rules listed here.
- Missing Pages
- Title missing on the front cover
- Incorrect pagination
- Books with typographical errors, such as misspellings or poor grammar.

Most new Project Gutenberg eBooks are in HTML format, which is validated using the W3C’s online validator. Additionally, the archive prefers plain text versions whenever possible due to their universal accessibility, and long-term usability. PDF, Word, and other word processor formats are avoided as master formats due to the challenges involved in converting them to HTML and updating them [@gutenberg].

With these points being addressed and the impartial nature of the archive being publicly disclosed on its website, the source data in which I garner my literary texts are selected with strict quality and accuracy. However, there are potential spaces for biases such as selection bias where Project Gutenberg primarily hosts older works that have entered the public domain. Thus, archived works might gear towards certain genres, authors, or time periods, potentially limiting the diversity of the datasets for analysis, which will be further discussed below [@gutenberg].

### Data Cleaning and Word Tokenization

```{r}
#| echo: false
#| eval: true
#| label: tbl-joyce
#| tbl-cap: "An Examplary Table Containing Unprocessed Novel Text (James Joyce)"
#| warning: false

portrait |>
  rename("Book ID" = gutenberg_id,
         "Text" = text,
         "Book" = book,
         "Author" = author) |>
  head(3) |>
  kable(
    col.names = c("Book ID", "Text", "Book", "Author"),
    booktabs = TRUE
  ) |>
  kable_styling(latex_options = c("striped", "hold_position"), 
                bootstrap_options = "striped", 
                full_width = FALSE)

```

To collect text data from five SOC authors using the `gutenbergr` package, I began by downloading the index of each author's works from the Project Gutenberg database. These authors include Joyce, Mansfield, Eliot, Woolf, and Proust [@rGuten] ([Appendix]).

Next, I created custom datasets containing each author's raw texts. This involved using the identification number of the desired literary work and the `gutenberg_download` function to retrieve the HTML texts. Subsequently, I saved the data into individual CSV files and later converted them into Parquet datasets, as demonstrated in @tbl-joyce example for James Joyce.

\newpage

```{r}
#| echo: false
#| eval: true
#| label: tbl-joyce-token
#| tbl-cap: Tokenized Novel Text & Word Count after Tokenization (James Joyce)
#| tbl-subcap: ["Tokenized Text", "Word Count"]
#| layout-ncol: 2
#| warning: false
#| message: false

portrait_clean <- portrait_clean |>
  select(-book)

portrait_clean |>
  rename("Book ID" = gutenberg_id,
         "Author" = author,
         "Word" = word) |>
  tail(3) |>
  kable(
    col.names = c("Book ID", "Author", "Word"),
    booktabs = TRUE
  ) |>
  kable_styling(latex_options = c("striped", "hold_position"), 
                bootstrap_options = "striped", 
                full_width = FALSE)

portrait_words |>
  rename("Word" = word) |>
  head(5) |>
  kable(
    col.names = c("Word", "Count"),
    booktabs = TRUE
  ) |>
  kable_styling(latex_options = c("striped", "hold_position"), 
                bootstrap_options = "striped", 
                full_width = FALSE)

```

Next, observable in @tbl-joyce-token, to measure the word count of each literary text, I first tokenized or separated the texts into individual chunks of words using `unnest_tokens()` [@rTidytext].In addition, to account for stop words like, a, of, and, etc., I leveraged the dataset `stop_words`and the function `anti_join` from @rTidytext to exclude all stop words from my self-generated datasets of literary texts, ensuring data validity and meaningful analysis.

Then, to quantify the number of times each word appears in each text, I used `count()` (@tbl-joyce-token). The result manifests into @fig-count.

\newpage
```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-count
#| layout-ncol: 2
#| fig-cap: "Comparative Analysis of Top 20 Word Frequencies from Famous SOC Authors"
#| fig-subcap: ["James Joyce", "Virginia Woolf", "Marcel Proust", "T.S. Eliot", "Katherine Mansfield"]

portrait_words |> head(20) |>
ggplot(aes(x = n, y = reorder(word, n))) +
  geom_bar(stat = "identity", fill = "steelblue4") +
  geom_text(aes(label = n), hjust = -0.3, vjust = 0.5, color = "grey40", size = 3) +
  labs(x = "Frequency", y = "Word")

dalloway_words |> head(20) |>
ggplot(aes(x = n, y = reorder(word, n))) +
  geom_bar(stat = "identity", fill = "steelblue4") +
  geom_text(aes(label = n), hjust = -0.3, vjust = 0.5, color = "grey40", size = 3) +
  labs(x = "Frequency", y = "Word")

swann_words |> head(20) |>
ggplot(aes(x = n, y = reorder(word, n))) +
  geom_bar(stat = "identity", fill = "steelblue4") +
  geom_text(aes(label = n), hjust = -0.3, vjust = 0.5, color = "grey40", size = 3) +
  labs(x = "Frequency", y = "Word")

prufrock_words |> head(20) |>
ggplot(aes(x = n, y = reorder(word, n))) +
  geom_bar(stat = "identity", fill = "steelblue4") +
  geom_text(aes(label = n), hjust = -0.3, vjust = 0.5, color = "grey40", size = 3) +
  labs(x = "Frequency", y = "Word")

bliss_words |> head(20) |>
ggplot(aes(x = n, y = reorder(word, n))) +
  geom_bar(stat = "identity", fill = "steelblue4") +
  geom_text(aes(label = n), hjust = -0.3, vjust = 0.5, color = "grey40", size = 3) +
  labs(x = "Frequency", y = "Word")

```

### Comparative Word Frequency

```{r}
#| echo: false
#| eval: true
#| label: tbl-frequency
#| tbl-cap: "Word Frequency of Stream of Consciousness Novels, A Comparison Between Five Authors "
#| warning: false

frequency <- bind_rows(mutate(portrait_words, author = "James Joyce"),
                       mutate(swann_words, author = "Marcel Proust"),
                       mutate(dalloway_words, author = "Virginia Woolf"),
                       mutate(bliss_words, author = "Katherine Mansfield"),
                       mutate(prufrock_words, author = "T.S. Eliot")) |>
  mutate(word = str_extract(word, "[a-z']+")) |>
  count(author, word) |>
  group_by(author) |>
  mutate(proportion = n / sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = author, values_from = proportion) |>
  filter(if_all(everything(), ~is.na(.) | !is.na(.)))

#reshape the frequency datasets and add sentiment values
bing_lexicon <- get_sentiments("bing") %>%
  rename(sentiment_binary = sentiment)

frequency_long <- pivot_longer(frequency, 
                               cols = c("James Joyce", "Marcel Proust", "Virginia Woolf", "Katherine Mansfield", "T.S. Eliot"), 
                               names_to = "author", values_to = "frequency") |>
  inner_join(bing_lexicon) |>
  inner_join(get_sentiments("afinn")) |>
  inner_join(get_sentiments("nrc"))

frequency_long |> 
  rename("Word" = word) |>
  head(5) |>
  kable(
    col.names = c("Word", "Author", "Frequency Percent", "Binary Sentiment", "Value", "Sentiment"),
    booktabs = TRUE
  ) |>  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
```

To accurately compare word count among SOC authors, from @tbl-frequency, I convert each word count from each author into a percentage. This is done by using `mutate` to calculate the word frequency of each word relative to the total word count of each author. Due to the different number of words present in each author's literary works, this method accounts for the inconsistencies of between each author's word count. As explained, the variables within @tbl-frequency are:

- **Word:** This variable is the word sourced from SOC literary texts.
- **Author:** This variable accounts for the author whose books contain the word in question.
- **Frequency:** This variable is the word count percentage of the word in question against each author's total word count.
- **Binary Sentiment (Bing):** This variable, as explained, comes from the Bing sentiment index, enabling us to analyze how every word is usually perceived emotionally, whether negatively or positively.
- **Value of Sentiment (AFINN):** This variable, as explained, comes from the AFINN sentiment index, enabling us to analyze how every word is usually perceived emotionally, on a scale between -5 and 5.
- **Sentiment (NRC):** This variable, as explained, comes from the NRC sentiment index, enabling us to analyze how every word is usually perceived emotionally, whether it's associated with anger, anticipation, disgust, fear, joy, negative, positive, sadness, surprise or trust.

\newpage
```{r}
#| echo: false
#| eval: true
#| label: tbl-boxplot 
#| tbl-cap: "Overall Boxplot Summary Statistics of all Word Frequency Percentages (AFINN Sentiment)"
#| warning: false

# Calculate summary statistics
frequency_long |>
  group_by(frequency) %>%
  summarise(
    min = min(value),
    q1 = quantile(value, probs = 0.25),
    median = median(value),
    mean = mean(value),
    q3 = quantile(value, probs = 0.75),
    max = max(value),
    sd = sd(value)
  ) |>
    head(8) |>
  kable(
    col.names = c("Frequency", "Min", "Q1", "Median", "Mean", "Q3", "Max", "SD"),
    booktabs = TRUE
  ) |>
  kable_styling(latex_options = c("striped", "hold_position"), 
                bootstrap_options = "striped", 
                full_width = FALSE)

```

@tbl-boxplot captures the text data garnered and their emotional connotations. Overseeing the entire dataset, @tbl-boxplot in particular demonstrates that, for all distinct word frequency, the median sentiment value is -2 on the scale between -5 and 5. More specifically, words that appear most frequently, or words with the frequency  of 0.0008511%, have the lowest mean sentiment value-or the most negative connotation of -2.33 on the scale between -5 and 5 in the AFINN sentiment index. This summary statistics further shows how SOC literature mostly associates with words that carry more negative or even extreme connotations often associated with mental health themes, and topics such as depression, disassociation or schizophrenia [@pathology], a central argument that I'll further discuss in [Results].

## Data Limitations

### Lack of Thorough Word Cleaning

With the large amount of different literary texts need processing by myself, it's hard to truly conduct thorough word processing. Certain texts, while can have their frequencies accounted for, can't have sentiment analysis due to their factual nature. These textual elements might include things like chapter titles, Roman numbers and personal names, which can strongly skew and affect the data integrity. As observed in @fig-count, "Swann", "Stephen" or "Jacob" are the most repeated words in Proust's, Joyce's and Woolf's, proving this data limitation.

### Potential Measurement Errors

The actual raw text data sourced from Project Gutenberg may be susceptible to measurement errors due to various factors, such as selection bias towards older, Western and canonical texts, social desirability bias that edits out certain political elements, memory lapses, or misinterpretations of literary texts by voluntary proofreaders. While the archive has strict guidelines to minimize these errors, they cannot be entirely eliminated.

### Uneven Text Length

I chose literary texts based on their global reception and canonical status within the stream-of-consciousness genre. This decision involved forgoing certain practical considerations, such as ensuring similar text lengths across all works by each author. While this approach enables a thorough exploration of word frequency analysis, sentiment value, and word networks within the SOC genre, it also has the potential to emphasize the frequencies of certain words from specific authors more than others due to the longer text length of those authors. This could critically affect the integrity of the data analysis.

### Biased Author Selection

Furthermore, SOC literature is expansive, spanning a wide range of themes, styles, and narrative techniques. By exclusively representing the genre through a select group of canonical Western authors, I risk constraining the genre's diversity. By oversimplifying the complexity of the the dynamic literary genre, I also affect the integrity of the datasets and its correlation to mental health themes.

# Results

## The Dominant Vocabulary of SOC Literature, A Word Frequency Analysis

@fig-count demonstrates that beside personal names, words with the highest counts are mostly abstract nouns that revolve around the nature of time, and life. For instance, these words are, \textit{god} (appearing 194 times in Joyce's), \textit{time} (appearing in 4 authors' texts, 121 times for Joyce's, 77 times for Woolf's, 421 times for Proust's, 23 times for Eliot), \textit{mind} (108 times in Joyce's and 196 times in Proust's), \textit{life} (appearing 266 times in Proust's and 131 times in Joyce's), etc.  This further exemplifies the sense of disassociation, temporality and non-linearity commonly associated to stream-of-consciousness literature and more broadly with mental health themes [@flow] [@pathology] [@resampling].

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-sentiment-frequency
#| layout-ncol: 2
#| fig-cap: "Overall Word Frequency and Sentiment Analysis"
#| fig-subcap: ["Bing lexicon", "NRC lexicon"]

# Reshape data for plotting
bing_lexicon <- get_sentiments("bing") %>%
  rename(sentiment_binary = sentiment)

frequency_long <- pivot_longer(frequency, 
                               cols = c("James Joyce", "Marcel Proust", "Virginia Woolf", "Katherine Mansfield", "T.S. Eliot"), 
                               names_to = "author", values_to = "frequency") |>
  inner_join(bing_lexicon) |>
  inner_join(get_sentiments("afinn")) |>
  inner_join(get_sentiments("nrc"))

frequency_long <- na.omit(frequency_long)

ggplot(frequency_long, aes(x = frequency, y = value, fill = sentiment_binary)) +
  geom_boxplot() +
  labs(title = "Boxplot of Sentiment Scores by Author (Bing)",
       x = "Word Frequency Percentage",
       y = "Sentiment Value (AFINN)",
       fill = "Sentiment") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("lightblue", "darkblue"))

ggplot(frequency_long, aes(x = frequency, y = value, fill = sentiment)) +
  geom_boxplot() +
  labs(title = "Boxplot of Sentiment Scores by Author (NRC)",
       x = "Word Frequency Percentage",
       y = "Sentiment Value (AFINN)",
       fill = "Sentiment") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c(
    "negative" = "darkblue", "sadness" = "steelblue4", "anger" = "royalblue4",
    "disgust" = "cadetblue", "fear" = "blue4", "joy" = "skyblue",
    "trust" = "lightblue", "anticipation" = "deepskyblue", "positive" = "dodgerblue",
    "surprise" = "lightsteelblue"
  ))

```

Within @fig-sentiment-frequency, for the NRC index, between all authors, the words with lower frequency-those ranging from 0.0000819% to 0004255%-are largely words that carry a mixture of emotional connotations such as, joy, surprise, positive, disgust and anticipation. However, words with high frequency-ranging from 0,0004255% to 0.0008511%-mostly contain negative connotations like fear and negative They also have a -1.875 median on a scale between -5 to 5 in the AFINN index. Similarly, for the Bing index, words with higher frequency-those ranging from 0,0004255% to 0.0008511%-largely carry negative connotations. They also have, approximately, a -1.875 median on a scale between -5 to 5 in the AFINN index. This further emphasizes how SOC literature mostly contains negativity-inclined vocabulary and linguistic patterns

To further explore the contextual relationship between each word, @fig-word-networks creates word networks of all nine of these SOC texts. These networks will capture how often each word frequents in these SOC texts and how closely each correlates with others, whether as single words (unigrams), pairs of words (bigrams), or triplets of words (trigrams). In particular, the darker the connecting lines, the more correlative these combinations of words are, and the bluer the words, the more frequent they appear.

 providing us with a thematic understanding of SOC literature.

\newpage
```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false

## Code sourced from https://www.youtube.com/watch?v=ae_XVhjHd_o

##first, n >=250
generate_word_graph <- function(combined_text,
                                minimum_n = 100,
                                minimum_correlation = 0.2) {

combined_text_count <- combined_text |>
  count(word, sort = TRUE) |>
  filter(n >= 150)

combined_correlations <- combined_text |>
  semi_join(combined_text_count, by = "word") |>
  pairwise_cor(item = word, feature = book) |>
  filter(correlation >= 1)

graph_from_data_frame(d = combined_correlations, #create graphs of networks of words
                      vertices = combined_text_count |>
                      semi_join(combined_correlations, by = c("word" = "item1"))) |>
  ggraph(layout = "fr") +
  geom_edge_link(aes(alpha = correlation), alpha = 0.05) +
  geom_node_point() +
  geom_node_text(aes(color = n, label = name), repel = TRUE)
}

##second, n >=300
generate_word_graph_second <- function(combined_text,
                                minimum_n = 100,
                                minimum_correlation = 0.2) {

combined_text_count <- combined_text |>
  count(word, sort = TRUE) |>
  filter(n >= 250)

combined_correlations <- combined_text |>
  semi_join(combined_text_count, by = "word") |>
  pairwise_cor(item = word, feature = book) |>
  filter(correlation >= 1)

graph_from_data_frame(d = combined_correlations, #create graphs of networks of words
                      vertices = combined_text_count |>
                      semi_join(combined_correlations, by = c("word" = "item1"))) |>
  ggraph(layout = "fr") +
  geom_edge_link(aes(alpha = correlation), alpha = 0.05) +
  geom_node_point() +
  geom_node_text(aes(color = n, label = name), repel = TRUE)
}

##third, n >=350
generate_word_graph_third <- function(combined_text,
                                minimum_n = 100,
                                minimum_correlation = 0.2) {

combined_text_count <- combined_text |>
  count(word, sort = TRUE) |>
  filter(n >= 350)

combined_correlations <- combined_text |>
  semi_join(combined_text_count, by = "word") |>
  pairwise_cor(item = word, feature = book) |>
  filter(correlation >= 0.6)

graph_from_data_frame(d = combined_correlations, #create graphs of networks of words
                      vertices = combined_text_count |>
                      semi_join(combined_correlations, by = c("word" = "item1"))) |>
  ggraph(layout = "fr") +
  geom_edge_link(aes(alpha = correlation), alpha = 0.05) +
  geom_node_point() +
  geom_node_text(aes(color = n, label = name), repel = TRUE)
}

## fourth, n>= 400
generate_word_graph_final <- function(combined_text,
                                minimum_n = 100,
                                minimum_correlation = 0.2) {

combined_text_count <- combined_text |>
  count(word, sort = TRUE) |>
  filter(n >= 400)

combined_correlations <- combined_text |>
  semi_join(combined_text_count, by = "word") |>
  pairwise_cor(item = word, feature = book) |>
  filter(correlation >= 0.8)

graph_from_data_frame(d = combined_correlations, #create graphs of networks of words
                      vertices = combined_text_count |>
                      semi_join(combined_correlations, by = c("word" = "item1"))) |>
  ggraph(layout = "fr") +
  geom_edge_link(aes(alpha = correlation), alpha = 0.05) +
  geom_node_point() +
  geom_node_text(aes(color = n, label = name), repel = TRUE)
}

# Code sourced from
# filtered_text <- combined_text[combined_text$n >= 100, ] #only select words with a frequency larger or equal to 100

# wordcloud(words = filtered_text$word, freq = filtered_text$n, scale=c(3,0.5), min.freq = 2, random.order = FALSE, colors= colorRampPalette(c("lightgray", "black"))(9))

```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-word-networks
#| layout-ncol: 2
#| fig-cap: "Word Networks Measured by Frequency and Correlation when Combining All Stream of Consciousness Novels"
#| fig-subcap: ["100 Frequency, 0.2 Correlation", "200 Frequency, 0.4 Correlation", "200 Frequency, 0.8 Correlation", "400 Frequency, 1 Correlation"]

combined_text|>
  generate_word_graph(minimum_n = 250, minimum_correlation = 0.2)

combined_text|>
  generate_word_graph_second(minimum_n = 300, minimum_correlation = 0.4)

combined_text|>
  generate_word_graph_third(minimum_n = 350, minimum_correlation = 0.6)

combined_text|>
  generate_word_graph_final(minimum_n = 400, minimum_correlation = 0.8)
```

## Sentiment Analysis

To further understand the emotional landscape and mental health themes related to SOC literature, I've combined all nine SOC texts to conduct an overall sentiment analysis [@sentiment].

For Bing index, @fig-sentiment-category shows that, SOC authors' literary works are predominantly negative. All authors, except for Mansfield, demonstrates this. 68% of Joyce's works are negative compared to his 32% positive, similar to Eliot's, which is also the highest statistic out of all five authors. Woolf comes second with 60% words carrying negative meanings, while 40% are negative. Proust comes fourth with  57% negative and 43% positive. While Mansfield has words with positive connotations, at 51%, this isn't much different from her 49% negative words.

The NRC index in @fig-sentiment-category further details the emotional nuances of SOC literature. The dominant emotional connotation in SOC literature's vocabulary is negative, standing at 16% to 22%. While the second highest percentage range belong to words with positive connotation, ranging from 11% to 17%, most of the other bars are occupied by words with negative-leaning emotions, such as sadness, fear, and anger. Sadness-related words, in particular, ranges the third highest, from 11% to 15%.

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-sentiment-category
#| layout-ncol: 2
#| fig-cap: "Categorical Sentiment Analysis of All SOC Novel Texts by Authors"
#| fig-subcap: ["Bing lexicon", "NRC lexicon"]

bing_lexicon <- get_sentiments("bing") |>
  rename(sentiment_binary = sentiment)

sentiment_combined <- combined_text |>
  inner_join(bing_lexicon) |>
  inner_join(get_sentiments("afinn")) |>
  inner_join(get_sentiments("nrc"))

sentiment_category_binary <- sentiment_combined |>
  group_by(author, sentiment_binary) |>
  count() |>
  group_by(author) |>
  mutate(percentage = round(n / sum(n) * 100))

# Plot sentiment values
ggplot(sentiment_category_binary, aes(x = author, fill = sentiment_binary, y = percentage)) +
  geom_bar(stat = "identity") +
    geom_text(aes(label = percentage),position = position_stack(vjust = 0.5), color = "white", size = 3) +
  labs(
       x = "Author", y = "Percentage", fill = "Sentiment") +
  scale_fill_manual(values = c("positive" = "lightblue", "negative" = "darkblue")) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),  legend.key.size = unit(0.5, "cm"))

sentiment_category <- sentiment_combined |>
  group_by(author, sentiment) |>
  count() |>
  group_by(author) |>
  mutate(percentage = round(n / sum(n) * 100))  |>
  group_by(author) |>
  arrange(desc(percentage))

# Plot sentiment values
ggplot(sentiment_category, aes(x = author, fill = sentiment, y = percentage)) +
  geom_bar(stat = "identity") +
   geom_text(aes(label = percentage),  position = position_stack(vjust = 0.5), color = "white", size = 3) +
  labs(
       x = "Author", y = "Percentage", fill = "Sentiment") +
   scale_fill_manual(values = c(
    "negative" = "darkblue", "sadness" = "steelblue4", "anger" = "royalblue4",
    "disgust" = "cadetblue", "fear" = "blue4", "joy" = "skyblue",
    "trust" = "lightblue", "anticipation" = "deepskyblue", "positive" = "dodgerblue",
    "surprise" = "lightsteelblue"
  )) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
       legend.key.size = unit(0.5, "cm"))

```

The AFINN index in @fig-sentiment-number portrays a similar trend with its more quantitative sentiment analysis. On a scale of -5 to 5, SOC literature from these five authors mostly carry a negative sentiment, observable in the significant difference in height between the negative trendlines and the positive trendlines. Eliot's works are the most negative with 37% of words being -2 on the emotional scale between -5 and 5. Respectively, Woolf, Proust and Joyce tail beind at 33%, 32% and 31%, averaging out to be ~-2.2 on the emotional scale between -5 and 5. While most a large percentage Mansfield's works are positive, her highest percentage is negative on the AFINN index, at 24% for a scale of about -2.2 between -5 and 5.

This further suggests that the SOC literary landscape is inundated with negative, disgust, fear and sadness-leaning vocabulary. These trends reflect the incoherent mental spaces, depressive perceptions, and existential angst inherent to SOC literature. These are the central themes often conveyed by SOC authors through their exploration of the unconscious mind and repressed emotions, reflecting the societal concerns of the late 19th to mid-20th century West [@pathology] [@resampling].

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-sentiment-number
#| fig-cap: "Numerical Sentiment Analysis of All SOC Novel Texts by Authors (AFINN lexicon)"

sentiment_value <- sentiment_combined |>
  inner_join(get_sentiments("bing")) |>
  group_by(author, value, sentiment) |>
  count() |>
  group_by(author) |>
  mutate(percentage = round(n / sum(n) * 100))

ggplot(sentiment_value, aes(x = value, y = percentage, color = sentiment, group = sentiment)) +
  geom_line() +
    geom_text(aes(label = percentage),position = position_stack(vjust = 0.5), color = "black", size = 2) +
  facet_wrap(~ author, scales = "free") +
  labs(
       x = "Sentiment Value", y = "Percentage") +
  scale_color_manual(values = c("positive" = "lightblue", "negative" = "darkblue")) + # Adjust line colors
  theme_minimal()
```

## Gendered Mental Landscape of Stream of Consciousness Novels

\newpage
```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-word-gender
#| layout-ncol: 2
#| fig-cap: "Comparative Analysis of Word Frequency in Female and Male Stream of Consciousness Authors"
#| fig-subcap: ["Female Authors", "Male Authors"]

# Calculate the word frequency over all unique words and bind all frequency rate of authors together to conduct a comparative comparison
# Code referenced from https://medium.com/the-data-nudge/nlp-basics-exploring-word-frequency-with-the-tidytext-r-package-ac35a6d805f4

frequency_female <- bind_rows(mutate(dalloway_words, author = "Virginia Woolf"),
                              mutate(bliss_words, author = "Katherine Mansfield")) |>
  mutate(word = str_extract(word, "[a-z']+")) |>
  count(author, word) |>
  group_by(author) |>
  mutate(proportion = n / sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = author, values_from = proportion) |>
  pivot_longer(`Virginia Woolf`,
               names_to = "author", values_to = "proportion")

ggplot(frequency_female, aes(x = proportion, y = `Katherine Mansfield`, color = abs(`Katherine Mansfield` - proportion))) +
  geom_abline(color = "gray50", lty = 2) +
  geom_jitter(alpha = 0.1, size = 1, width = 0.5, height = 0.5) +  
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = scales::percent_format()) +
  scale_y_log10(labels = scales::percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), 
                       low = "lightblue", high = "darkblue") +
  facet_wrap(~author, ncol = 1) +
  theme(legend.position="none") +
  labs(y = "Katherine Mansfield", x = NULL)

frequency_male <- bind_rows(
                       mutate(swann_words, author = "Marcel Proust"),
                       mutate(prufrock_words, author = "T.S. Eliot"),
                       mutate(portrait_words, author = "James Joyce")) |>
  mutate(word = str_extract(word, "[a-z']+")) |>
  count(author, word) |>
  group_by(author) |>
  mutate(proportion = n / sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = author, values_from = proportion) |>
  pivot_longer(`Marcel Proust`:`James Joyce`,
               names_to = "author", values_to = "proportion")

ggplot(frequency_male, aes(x = proportion, y = `T.S. Eliot`, color = abs(`T.S. Eliot` - proportion))) +
  geom_abline(color = "gray50", lty = 2) +
  geom_jitter(alpha = 0.1, size = 1.5, width = 0.8, height = 0.8) +  
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), 
                       low = "lightblue", high = "darkblue") +
  facet_wrap(~author, ncol = 2) +
  theme(legend.position="none") +
  labs(y = "T.S. Eliot", x = NULL)

```

### Comparing SOC Literature's Female Authors

### Comparing SOC Literature's Male Authors

## Transnational SOC Novels and Mental Health Themes

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-frequency-int
#| fig-cap: "Comparative Analysis of Word Frequency in Transnational Stream of Consciousness Authors"

frequency_international <- bind_rows(
                       mutate(swann_words, author = "Marcel Proust"),
                       mutate(bliss_words, author = "Katherine Mansfield"),
                       mutate(portrait_words, author = "James Joyce")) |>
  mutate(word = str_extract(word, "[a-z']+")) |>
  count(author, word) |>
  group_by(author) |>
  mutate(proportion = n / sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = author, values_from = proportion) |>
  pivot_longer(`Katherine Mansfield`:`James Joyce`,
               names_to = "author", values_to = "proportion")

ggplot(frequency_international, aes(x = proportion, y = `Marcel Proust`, color = abs(`Marcel Proust` - proportion))) +
  geom_abline(color = "gray50", lty = 2) +
  geom_jitter(alpha = 0.05, size = 1, width = 0.8, height = 0.8) +  
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), 
                       low = "lightblue", high = "darkblue") +
  facet_wrap(~author, ncol = 1) +
  theme(legend.position="none") +
  labs(y = "Marcel Proust", x = NULL)

```

# Discussion

## Mental Health Vocabulary: Patterns and Trends

Discuss vocabulary patterns and word trends

## Insights into Socio-Political Landscape of the West's Modernist Era

pathological consequences of the constant
shifts of identity in migrant characters. Most postcolonial studies contend that
marginalization occasioned by the dominant group result in neurotic conditions in
members of the minority group.

The novels' linguistic patterns reflect the socio-political landscape of the Western hemisphere, from  late 19th century to the mid-20th century.

Stream of consciousness literature emerged as a significant literary technique in the early 20th century, particularly associated with the modernist movement in literature. Several factors contributed to its rise and prominence:

Psychological Realism: One of the primary motivations for the development of stream of consciousness was the desire to represent the inner workings of the human mind more accurately. Writers sought to capture the complexity of thought processes, emotions, and perceptions in a more realistic and nuanced manner.
Response to Realism and Naturalism: Stream of consciousness literature can be seen as a reaction against the more structured and external focus of realism and naturalism. Instead of portraying characters and events objectively, authors wanted to delve deeper into subjective experiences and individual consciousness.
Technological and Cultural Changes: The early 20th century was a period of rapid technological advancement and social change, including the rise of psychology as a discipline. These developments influenced writers to explore new narrative techniques and modes of representation.
Literary Experimentation: Modernist writers were often interested in pushing the boundaries of traditional narrative forms and experimenting with language, style, and structure. Stream of consciousness offered a way to challenge conventional storytelling techniques and engage readers in more innovative ways.
Exploration of Identity and Perception: Stream of consciousness literature allowed writers to explore themes related to identity, consciousness, and perception. By immersing readers in the internal thoughts and sensations of characters, authors could interrogate questions about the nature of reality, self-awareness, and the construction of meaning.
Overall, stream of consciousness literature became a prominent feature of modernist writing due to its capacity to convey the complexity of human experience and its ability to reflect the uncertainties and ambiguities of the modern world.

The rise of modernism in literature, art, and culture during the late 19th and early 20th centuries was influenced by several interconnected factors:

Industrialization and Urbanization: The rapid industrialization and urbanization of society brought about significant changes in the way people lived, worked, and interacted. This upheaval in traditional social structures and lifestyles led to a sense of dislocation and fragmentation, which modernist artists and writers sought to capture in their work.
Technological Advances: Advances in technology, particularly in transportation, communication, and mass media, reshaped human experience and perception of time and space. Innovations such as the telegraph, telephone, and photography altered how people communicated, experienced distance, and understood reality.
Crisis of Faith and Meaning: The late 19th and early 20th centuries witnessed a crisis of faith and meaning, as traditional religious and philosophical beliefs came into question. Developments such as Darwin's theory of evolution, Freudian psychology, and existential philosophy challenged established notions of human existence, morality, and purpose.
World Wars and Political Turmoil: The turmoil of World War I and the interwar period brought about profound changes in society, politics, and culture. The devastation of war, the collapse of empires, and the emergence of new ideologies (such as communism and fascism) shattered old certainties and created an atmosphere of disillusionment and uncertainty.
Intellectual and Cultural Exchange: The modernist movement was characterized by a spirit of intellectual and cultural exchange, with artists and writers drawing inspiration from diverse sources, including non-Western cultures, folk traditions, and avant-garde movements. This openness to new ideas and influences fostered experimentation and innovation in the arts.
Quest for Individuality and Authenticity: Modernism rejected the conventions and constraints of traditional artistic forms and sought to express individual subjectivity and experience in new and unconventional ways. Artists and writers embraced themes of alienation, identity, and the search for authenticity in an increasingly fragmented and alienating world.

## Schizophrenic and Disassociative Tendencies in Female Stream of Consciousness

## Weaknesses

### Lack of Thorough Word Cleaning

This includes words such as chapter titles, Roman numbers and personal names, affecting the integrity of data analysis ("Swann" being the most common word)

### Decontextualized Literature Works and Limiting Publication Editions

Words are singled out and analyzed without context which could have affected their intended meanings, especially in such a complex genre such as stream of consciousness. The limiting novel editions also doesn't make sure that their literary integrity are maintained and the analysis might have missed important texts of other editions.

While these datasets aim to illuminate on the general time trends of childlessness with each variable measured, they are often de-contextualized for the same purpose. For instance, in Table 6, there is no clarifying note on whether people who declared themselves as ‘divorced’ are divorced once or more. This can be extremely narrowing, and exclude nuances when analyzing the contexts behind childlessness.

### Reliance on Three Different Indices

Since the novels are chosen based on their worldwide reception and canon, practical constraints like how all novels should be of similar length are ignored. This constraint can prevent one novel having more words than others, which can critically affect the integrity of the word frequency analysis, sentiment value and word networks where one novel dominates the others and skews the results.

In addition, these authors work are expansive and so choosing a select few to bind them to the SOC genre can be limiting as the SOC genre in itself is already an amalgamation of different literary trends. Thus, this can affect the integrity of the datasets.

## Moving Forward and Next Steps

\newpage

# Appendix

## Additional Data Details

```{r}
##| eval: true
##| echo: false
##| message: false
##| warning: false
#combined_books 
```

### Data Gathering

```{r fig.pos="H"}
##| echo: false
##| message: false
##| label: tbl-reasons-strip-search
##| tbl-cap:

```

### Data Cleaning

```{r fig.pos="H"}
##| echo: false
##| message: false
##| label: tbl-items-strip-search
##| tbl-cap:

```

\newpage

# References